---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true                
    toc_float: true 
    code_folding: "hide"       
---

```{r, include = FALSE}
library(tidyverse)
```

### Data Description
```{r, message = FALSE} 
final_long=
  read_csv("Data/final_long.csv")

final_long %>%
  select(admission_prop, maskwearing_prop, symptom_prop,  finances_prop, depression_prop) %>%
  summary() %>%
  knitr::kable(digits = 3)
```

Our data has `r nrow(final_long)` observations and `r ncol(final_long)` variables and tells us about maskwearing, influenza symptoms, and admissions across the U.S. from years `r range(final_long$date)[1]` to `r range(final_long$date)[2]`. The average percentage of people who say they wear a mask most or all of the time when they are in public in the past 7 days is 52.75%. The average percentage of adult and pediatric hospital admissions with confirmed influenza occurring each day (7-day average) is 0.05%, the average percentage of people self-reporting influenza-like symptoms such as fever with cough or sore throat in the past 7 days is 1.81%, the average percentage of people  who reported feeling depressed for most or all of the past 7 days is 12.61%, and the average percentage of people who report being very or somewhat worried about their "household's finances for the next month" is 37.45%.

### Overall Trends in the U.S.for Participants Wearing Masks between 2021 - 2022
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_maskwearing_prop = mean(as.numeric(maskwearing_prop))) %>%
  ggplot(aes(x = date, y = avg_maskwearing_prop)) +
  geom_line()+
  labs(title = "Average Percentage of People Wearing Masks", x = "Date", y = "Percentage of People Wearing Masks")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```
Based on the graph, we can see that there was a huge drop from May 2021 to July 2021. This was due to the fact that in May 2021, as the United States continued to see dramatic improvements in COVID-19 cases, federal officials announced a sweeping relaxation of face mask guidelines allowing fully vaccinated people from not wearing masks in most places, either outdoors or inside (https://www.latimes.com/science/story/2021-07-27/timeline-cdc-mask-guidance-during-covid-19-pandemic). However, mask wearing rate increased around September 2021 potentially because of a continued uptick in positive COVID-19 cases due to the Delta variant. (https://abc7ny.com/nyc-face-mask-guidance-restaurants-vaccine-mandate-danny-meyer-delta-variant/10924328/). Lastly, we see another hugh drop starting in February 2022 as infections in the U.S. decreased by 90% from a pandemic high during, leading the states to lift the mask mandates (https://www.cnbc.com/2022/02/22/covid-infections-plummet-90percent-from-us-pandemic-high-states-lift-mask-mandates.html). 

### Overall Trends in the U.S. for Participants Wearing Masks between 2021 - 2022
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_symptom_prop = mean(as.numeric(symptom_prop))) %>%
  ggplot(aes(x = date, y = avg_symptom_prop)) +
  geom_line() +
  labs(title = "Average Percentage of Participants with Influenza Symptom", x = "Date", y = "Percentage of People with Influenza Symptom") +
  theme(plot.title=element_text(face="bold", color="blue")) +
  theme(axis.text.x = element_text(angle=80, hjust=1))
```

### Overall Trends in the U.S. for Participants Worried about Finance between 2021 - 2022
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_finances_prop = mean(as.numeric(finances_prop))) %>%
  ggplot(aes(x = date, y = avg_finances_prop)) +
  geom_line()+
  labs(title = "Average Percentage of Participants worried about finance", x = "Date", y = "Percentage of participants worried about finance")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```

### Overall Trends in the U.S. 2021-2022 for Depressed Participants 
```{r}
final_long %>%
  drop_na %>%
  group_by(date)%>%
  summarize(avg_depression_prop = mean(as.numeric(depression_prop))) %>%
  ggplot(aes(x = date, y = avg_depression_prop)) +
  geom_line()+
  labs(title = "Average Percentage of depressed participants", x = "Date", y = "Percentage of depressed participants")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```

### Overall Trends in the U.S. for Hospital Admissions with Confirmed Influenza between 2021 - 2022
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_admission_prop = mean(as.numeric(admission_prop))) %>%
  ggplot(aes(x = date, y = avg_admission_prop)) +
  geom_line()+
  labs(title = "Average Number of Hospital Admissions with Confirmed Influenza", x = "Date", y = "Number of Hospital Admissions")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```




