---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true                
    toc_float: true        
---

##Data Description
```{r, message=FALSE} 
final_long=
  read_csv("Data/final_long.csv")

final_long%>%
  select(admission_prop, maskwearing_prop, symptom_prop,  finances_prop, depression_prop) %>%
  summary()%>%
  knitr::kable(digits = 3)
```

Our data has `r nrow(final_long)` observations and `r ncol(final_long)` variables and tells us about maskwearing, influenza symptoms, and admissions across the U.S. from years `r range(final_long$date)[1]` to `r range(final_long$date)[2]`. The average percentage of people who say they wear a mask most or all of the time when they are in public in the past 7 days is 52.87%. The average number of adult and pediatric hospital admissions with confirmed influenza occurring each day (7-day average) is 131.54, and the average percentage of people self-reporting influenza-like symptoms such as fever with cough or sore throat in the past 7 days is 1.81%. 

### Overall Trends in the U.S. 2021-2022 for Participants Wearing Masks
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_maskwearing_prop = mean(as.numeric(maskwearing_prop))) %>%
  ggplot(aes(x = date, y = avg_maskwearing_prop)) +
  geom_line()+
  labs(title = "Average Percentage of People Wearing Masks", x = "Date", y = "Percentage of People Wearing Masks")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```

### Overall Trends in the U.S. 2021-2022 for Participants Wearing Masks
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_symptom_prop = mean(as.numeric(symptom_prop))) %>%
  ggplot(aes(x = date, y = avg_symptom_prop)) +
  geom_line() +
  labs(title = "Average Percentage of Participants with Influenza Symptom", x = "Date", y = "Percentage of People with Influenza Symptom") +
  theme(plot.title=element_text(face="bold", color="blue")) +
  theme(axis.text.x = element_text(angle=80, hjust=1))
```

### Overall Trends in the U.S. 2021-2022 for Participants worried about finance
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_finances_prop = mean(as.numeric(finances_prop))) %>%
  ggplot(aes(x = date, y = avg_finances_prop)) +
  geom_line()+
  labs(title = "Average Percentage of Participants worried about finance", x = "Date", y = "Percentage of participants worried about finance")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```
### Overall Trends in the U.S. 2021-2022 for Depressed Participants 
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_depression_prop = mean(as.numeric(depression_prop))) %>%
  ggplot(aes(x = date, y = avg_depression_prop)) +
  geom_line()+
  labs(title = "Average Percentage of depressed participants", x = "Date", y = "Percentage of depressed participants")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```

### Overall Trends in the U.S. 2021-2022 for Hospital Admissions with Confirmed Influenza
```{r}
final_long%>%
  drop_na%>%
  group_by(date)%>%
  summarize(avg_admission_prop = mean(as.numeric(admission_prop))) %>%
  ggplot(aes(x = date, y = avg_admission_prop)) +
  geom_line()+
  labs(title = "Average Number of Hospital Admissions with Confirmed Influenza", x = "Date", y = "Number of Hospital Admissions")+
  theme(plot.title=element_text(face="bold", color="blue"))+
  theme(axis.text.x = element_text(angle=80, hjust=1))
```




