---
title: "Model Building"
output: 
  html_document:
    code_folding: "hide"
---

```{r load_package, echo = FALSE, message=FALSE}
rm(list=ls())
library(tidyverse)
library(purrr)
library(geepack)
options(tibble.print_min = 5)
options(pillar.sigfig = 5)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

Model Selection
* Using a GEE model for the repeated measure of `maskwearing_prop`, `finances_prop`, and `depression_prop` at the state level, we were able to observe the associations between masking wearing rate and influenza hospital admission or experiencing flu-like symptoms. 

When evaluating the models at the national level, we used a backward-selection-like method. We started with the full model with all the covariates and interaction terms. Then we checked the p-value of all covariates and interaction terms with the outcome variable and found out that only`maskwearing_prop`, `finances_prop`, and `depression_prop` variables were significant, and none of the interaction terms were significant. Thus, we proceeded with the reduced model, only with `maskwearing_prop`, `finances_prop`, and `depression_prop` covariates. 

## Regression Analysis
```{r import_data}
final_long=
  read_csv("Data/final_long.csv")

states <- unique(final_long$state)
```

* excluded `pr, vi, as` since they do not have any measurements.

### Model Building

```{r}
analysis_df <- final_long %>% mutate(date=as.Date(date),state=as.factor(state)) %>% arrange(state,date) %>% select(everything())

# outcome = "admission" or "symptom"; l_date = lower bound of date written as "YYYY-MM-DD"; u_date = upper bound of date written as "YYYY-MM-DD"
# s_state = selected state; b_nation = model scale "nation" or "state" -wide

shiny_input_model <- function(outcome, l_date, u_date, s_state, s_cov, b_nation){
  cov_in<-as.character(s_cov)
  
  if (outcome=="admission"){
    model_formula <- formula(paste0("admission_prop ~ maskwearing_prop + ",cov_in))
  }else if (outcome=="symptom"){
    model_formula <- formula(paste0("symptom_prop ~ maskwearing_prop + ", cov_in))
  }
  
  if (b_nation=="state"){
    temp_df<-analysis_df %>% filter(state_full %in% s_state & date>=as.Date(l_date) & date<=as.Date(u_date))
    out_table <- geeglm(formula=model_formula, data=temp_df, id=state,
                        family=gaussian("identity"),corstr="exchangeable",std.err = "san.se") %>% 
      broom::tidy() %>% knitr::kable(digits = 3) 
  } else if (b_nation=="nation"){
    temp_df<-analysis_df %>% filter(date>=as.Date(l_date) & date<=as.Date(u_date))
    out_table <- geeglm(formula=model_formula, data=temp_df, id=state,
                        family=gaussian("identity"),corstr="exchangeable",std.err = "san.se") %>% 
      broom::tidy() %>% knitr::kable(digits = 3) 
  }

  return(out_table)
}

ex_1<-shiny_input_model(outcome="admission", l_date="2021-06-01", u_date="2022-04-01", s_state="Colorado", s_cov="finances_prop + depression_prop + finances_prop*depression_prop", b_nation="state")


ex_2<-shiny_input_model(outcome="admission", l_date="2021-06-01", u_date="2022-04-01", s_state=NA, s_cov="finances_prop + depression_prop + finances_prop*depression_prop", b_nation="nation")

```

* This model is a primitive speculative model, which we logically assume the two covariates had potentials to impact the associations. 

* Given exposure `maskwearing` is a continuous variable, we will be estimating the mean function of the repeated measure.

  * `identity` function is used as the link function

  * We assumed that the correlation among distinct observations in the same state is the same regardless of time. Therefore, we used `exchangeable` as the correlation structure.

  * `state` is treated as the `id`.

  * the function `shiny_input_model` was written in preparation for the shiny app deployment

* If we want to model the association in `co`, the function will return the following table: `r ex_1`

* If we want to model the association nationwide, the function will return the following table: `r ex_2`


### Related Final Report Section: Main Association

A Significant negative association between `maskwearing_prop` and `admission_prop`  (- 0.00151, P < 0.0001) was observed at the national level, adjusting for ‘finance_prop’ and ‘depression_prop’.
  
 * Since we cannot conclude causation from our model, we can only say that a higher mask-wearing rate was associated with lower hospital admissions due to severe influenza. 
  
  * This can be explained by the protective nature of masks. People who wore masks might have had a lower probability of contracting severe influenza infection, which could have been attributed to the reduced viral load due to mask filtration and less severe infection. 

A significant positive association between `maskwearing_prop` and `symptoms_prop` (0.01139, P < 0.0001) was observed at the national level, adjusting for ‘finance_prop’ and ‘depression_prop’.

  * Again since causation cannot be ascertained, we could only conclude that a higher mask-wearing rate was associated with higher flu-like symptoms. 
  
  * One possible explanation for this is that people may have become more aware of prevention for the flu if they were experiencing flu-like symptoms. Therefore, people might have tended to wear masks if they felt like they were sick.

The parameters for the predictors are seemingly small. However, in the context of the dataset where all the measurements were the percentages of the population, a small change could mean a huge variation clinically. The parameters for `admission_prop` are considerably smaller than those for `symptoms_prop` because the ceiling of `admission_prop` is much lower than that of `symptoms_prop`.

