---
title: "Model Building"
output: 
  html_document:
    code_folding: "hide"
---

```{r load_package, echo = FALSE, message=FALSE}
rm(list=ls())
library(tidyverse)
library(purrr)
library(geepack)
options(tibble.print_min = 5)
options(pillar.sigfig = 5)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

Model Selection
* Using a GEE model for the repeated measure of `maskwearing_prop`, `finances_prop`, and `depression_prop` at the state level, we were able to observe the associations between masking wearing rate and influenza hospital admission or experiencing flu-like symptoms. 

When evaluating the models at the national level, we used a backward-selection-like method. We started with the full model with all the covariates and interaction terms. Then we checked the p-value of all covariates and interaction terms with the outcome variable and found out that only`maskwearing_prop`, `finances_prop`, and `depression_prop` variables were significant, and none of the interaction terms were significant. Thus, we proceeded with the reduced model, only with `maskwearing_prop`, `finances_prop`, and `depression_prop` covariates. 

## Regression Analysis
```{r import_data}
final_long=
  read_csv("Data/final_long.csv")

states <- unique(final_long$state)
```

* excluded `pr, vi, as` since they do not have any measurements.

### Model Building

```{r}
analysis_df <- final_long %>% mutate(date=as.Date(date),state=as.factor(state)) %>% arrange(state,date) %>% select(everything())

# test <-analysis_df %>% filter(date>=as.Date("2021-06-01") & date<=as.Date("2022-04-01"))

# outcome = "admission" or "symptom"; l_date = lower bound of date written as "YYYY-MM-DD"; u_date = upper bound of date written as "YYYY-MM-DD"
# s_state = selected state; b_nation = model scale "nation" or "state" -wide

shiny_input_model <- function(outcome, l_date, u_date, s_state, s_cov, b_nation){
  cov_in<-as.character(s_cov)
  
  if (outcome=="admission"){
    model_formula <- formula(paste0("admission_prop ~ maskwearing_prop + ",cov_in))
  }else if (outcome=="symptom"){
    model_formula <- formula(paste0("symptom_prop ~ maskwearing_prop + ", cov_in))
  }
  
  if (b_nation=="state"){
    temp_df<-analysis_df %>% filter(state_full %in% s_state & date>=as.Date(l_date) & date<=as.Date(u_date))
    out_table <- geeglm(formula=model_formula, data=temp_df, id=state,
                        family=gaussian("identity"),corstr="exchangeable",std.err = "san.se") %>% 
      broom::tidy() %>% knitr::kable(digits = 3) 
  } else if (b_nation=="nation"){
    temp_df<-analysis_df %>% filter(date>=as.Date(l_date) & date<=as.Date(u_date))
    out_table <- geeglm(formula=model_formula, data=temp_df, id=state,
                        family=gaussian("identity"),corstr="exchangeable",std.err = "san.se") %>% 
      broom::tidy() %>% knitr::kable(digits = 3) 
  }

  return(out_table)
}

ex_1<-shiny_input_model(outcome="admission", l_date="2021-06-01", u_date="2022-04-01", s_state="Colorado", s_cov="finances_prop + depression_prop + finances_prop*depression_prop", b_nation="state")


ex_2<-shiny_input_model(outcome="admission", l_date="2021-06-01", u_date="2022-04-01", s_state=NA, s_cov="finances_prop + depression_prop + finances_prop*depression_prop", b_nation="nation")

```

* this model is a primitive speculative model, which we logically assume the two covariates had potentials to impact the associations. 

* given exposure `maskwearing` is a continuous variable, we will be estimating the mean function of the repeated measure.

  * `identity` function is used as the link function

  * We assumed that the correlation among distinct observations in the same state is the same regardless of time. Therefore, we used `exchangeable` as the correlation structure.

  * `state` is treated as the `id`.

  * the function `shiny_input_model` was written in preparation for the shiny app deployment

* If we want to model the association in `co`, the function will return the following table: `r ex_1`

* If we want to model the association nationwide, the function will return the following table: `r ex_2`


