---
title: "Final Project Proposal"
author: Mingee Choi(mlc2287), Barry Chiu (pc2979), Junyu Qi (jq2342), Soungbin Yim (sy3019), Troy Zhou (zz2901)
date: 11/11/2022
output: github_document
---

# Tentative project title
* The association between decrease in mask-wearing and the prevalence of seasonal influenza in 2022 

# Motivation for this project
* In the United States, the flu season usually occurs in the fall and winter, meaning people are most vulnerable to the flu during this season. However, as Covid restrictions and policies have become less stringent, we have seen a significant decrease in the number of people wearing masks. We wanted to explore how people wearing masks has an effect on disease prevalence.

* The results from Froses et al. paper showed that several combinations of mask efficacy and prevalence could substantially reduce the burden of seasonal influenza, suggesting that mask-wearing may be a viable alternative or complement to influenza vaccinations. Thus, we want to investigate whether the current trend of decrease in wearing masks is associated with the prevalence of influenza in 2022.  

  * Froese, H., & Prempeh, A. G. (2022). Mask use to curtail influenza in a post–COVID-19 world: modeling study. JMIRx Med, 3(2), e31955.

# Intended final products

* Website detailing motivation behind analysis, overview of data, dashboard visualizing mask-wearing trends and influenza prevalence over time, statistical analysis of association between mask wearing and influenza cases

# Anticipated data sources

* ‘Covidcast’ https://delphi.cmu.edu/covidcast/indicator/?date=20210324&sensor=fb-survey-smoothed_wwearing_mask_7d
* https://www.tableau.com/learn/articles/free-public-data-sets


# Planned analyses/visualizations/coding challenges

* Analysis

  * Logistic regression analysis with mask-wearing as the main exposure/predictor and the influenza cases will be the main outcome of interest.
Temporal comparison of influenza cases vs mask wearing trends between period of mask-mandate and period after mask-mandate

* Visualization

  * Sample Demographic Characteristics Table

  * Exploratory analysis of trends of the main outcome of interest, the influenza cases, by various covariates in NYC(e.g., race, age, etc) 

  * Interactive maps of mask wearing rate by various covariates in NYC (e.g., race, age, etc) 

* Coding challenges:

  * Data set cleaning will be messy since dealing with nationwide longitudinal data 
  
  * Missing data could also be challenging as some regions could be missing observations for some time periods
  
  * Merging datasets of the disease outcome with the masking rate dataset could be complex due to difficulties in finding appropriate matching variables 

# Planned timeline
* 11/15 - 11/18: project review meeting
* 11/19 - 11/23: data collection and clearing
* 11/24 - 11/30: data analysis & Graphics Output
* 12/1 - 12/5: producing Written Report
* 12/6 - 12/9: producing screencast video
* 12/10: Written report & webpage and screencast; peer assessment DUE
* 12/15: in class discussion

---
title: "AirBNB Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(flexdashboard)
library(plotly)
```

```{r, include = FALSE}
data("nyc_airbnb")
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(rating = review_scores_location / 2) %>% 
  select(
    borough = neighbourhood_group, neighbourhood, price, room_type,
    lat, long, rating) %>% 
  filter(
    borough == "Manhattan",
    room_type == "Entire home/apt",
    price %in% 100:500
  ) %>% 
  drop_na(rating)
```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}
nyc_airbnb %>% 
  mutate(
    text_label = str_c("Neighborhood: ", neighbourhood)
  ) %>% 
  plot_ly(
    x = ~lat, y = ~long, color = ~price,
    type = "scatter", mode = "markers",
    alpha = .5, text = ~text_label
  )
```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}
nyc_airbnb %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  plot_ly(
    y = ~price, color = ~neighbourhood,
    type = "box", colors = "viridis")
```

### Chart C

```{r}
nyc_airbnb %>% 
  count(neighbourhood) %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) %>%
  plot_ly(
    x = ~neighbourhood, y = ~n,
    type = "bar")
```
